\thispagestyle{empty}
\chapter*{Acknowledgement's}

The content here presented was produced during the last academic year of my bachelor's degree in Mathematics at the Institute of Mathematics and Statistics of the University of São Paulo (IME-USP). This material represents a very important end of cycle for me, and therefore, I need to thank those responsible for not letting me give up on everything.

First of all, I couldn't forget to mention my dear parents: \textsc{Ricardo Mezzovilla Gonçalves} and \textsc{Manoelina Pauzner}. They are the ones who know how passionate I am about mathematics, and in particular, my Mother has always shown the greatest interest, asking me to explain different and diverse topics. The opportunity to study was granted to me by them, in addition to always encouraging me. I must also not forget about my older sister: \textsc{Amanda Mezzovilla}, which had a hard time reviewing my gramatical errors, going through a heavy dose of math. Since I recognize myself as a ruminant human being, we had some astonishing philosophical discussions and great sharing of knowledge. Hence, all my family deserves some pizza slices.

During my studies, I also worked in parallel on completely orthogonal topics, to which I've been struggling to balance those bad rocks. Giving up was constantly on my mind. To the love of my life, \textsc{Maiara dos Anjos Justino}, I must thank for preventing me from giving up, giving me the strength to keep going, in a way only she would be able to lift me up. Only she saw my journeys against the tide, wearing me down not only physically but psychologically. Even if it's cliche, I don't know what I would do without her presence all these years. I hand her the biggest pizza slice possible (and even the olives that would be on my plate).

I must also thank my advisor \textsc{Severino Toscano do Rego Melo} who agreed to embark on this journey with me, in which I'd seen myself lost several times and, without his help, I wouldn't be able to continue. Speaking of pizza, he owes me and the boys who were his students in ``Differential Equations I'' at the second semester of 2019. So, I guess we're even on what comes to me.

I would like to make a few honorable mentions to some friends of mine, with the leftover pizza that was stale the next day\footnote{Do not forget that these are the one who tastes better!}. They were, directly or indirectly responsible for my journey up here: 
	\begin{itemize}
		\item \textsc{Francisco Felipe Chagas Rangel};
		\item \textsc{Zeus Santos Nazaro da Silva};\item \textsc{Isabella Matias Silva};
		\item \textsc{Guilherme Cerqueira Gonçalves}; \item \textsc{Andrey Luan Gomes Contel}.
	\end{itemize}
\thispagestyle{empty}
Fortunately, I have a large list of friends I know I can count on, and I would like to name all my friends from College, Work and the other four corners of the world, but I have a terrible memory and I do not want to be in anyone's debt for more pizzas.
